### 2008-09-17 Robust/Transactions/Notes/sysgurantees.tex
*** empty log message ***

The flow of program should be in a manner that we could have an arbitrary serial order of the members of $T$ (e.g. $T_5,T_6,T_1$,...) or any other order. The serial order should be in such a way that for any given two members of the set $T$, either the effect of all opertaions of \{ TransactionalFile.Read, TransactionalFileWrite, TransactionalFile.Seek()\} within $T_i$ should be visible to $T_j$ or none should be visible. In the first case we would say the $T_i$ precedes $T_j$ and in the seconds case $T_j$ preceds $T_i$ and $T_i$ can observe all the effects made by the operations in $T_j$. \\

@@

Def 1- 1-An operation in $T_i$ is said to have "changed" or "affected" some data structure in correspondece to all members of $OP_{T- T_i}$ , when it reflects the changes in the filesystem in such a way that the modification is durable and irreversible. A transaction with such an operation $T_i$ in its operation set is said to have changed the data 2- An operation in $T_i$ is said to have "changed or affeced the data in correspondence to all the operations in $OP_{T_i}$ as soon as the operation is encountered \\

Def 2- An operation in $T_i$ is said to have "used" the data when it has read some data and has done some actions that is potentially irreversible. hence if the data in question was to change, $op$ would have different results. A transaction with such an operation in its operations set is said to have used that data. Any data read and used within a transaction at its commit instant does not count as using. "Use" can just be attributed to the operations actually reading data before commit instant. \\

Def 3- A transaction $T_i$ is said to "commit" at instant $t_i$, if and only if it reflects its "changes" in the filesystem or equivalently makes it changes visible to the whole system for all members of $T$ accesing the data at instant $t_j$ such that $t_j > t_i$ . An operation $\in OP_{T_i}$ is said to commit if and only if $T_i$ commits. Commit instant are not splittable, hence it appears to other transactions that all changes are refleted in the file system together.\\

Def 4-  $\forall op_i \in OP_{T_i}$, $\forall op_j in OP{T_i}$, $op_i$ precedes $op_j$ if and only if $op_j$ does not use the data changed by $op_i$ OR if it does, then $op_j$ sees the changes made by $op_i$ unless data has been changed by a more recent precedessor. Furthermore  \\

Def 5- Succeds relationship is defined similarly.  \\

Def 6- If operation $a$ both precedes and succeds operation $b$ since they do not share any resources so the changes can be made visible, $a$ both precedes and succeds $b$. \\

Def 7- $\forall$ $op_{T_i}$ $\in OP_{T_i}$ and $\forall op_{T_j}\in OP_{T_j}$ if and only if $op_{T_i} \rightarrow op_{T_j}$,  then $T_i$ \emph {can} $\rightarrow T_j$) \hspace{8mm} (this defines precedes relationship for members of $T$) \\

Def 8- An operation $op \in OP_{T_i}$ should be able to see the "changes" 1- made by $OP_{T_commited}$ and 2-$oper \in OP_{T_i}$ that in natural order of transaction occur before $op$

\subsection{Rules}
If we denote $\rightarrow$ as the relation precedes between operations and $OP_{T_i}$ as the set of all operation within $T_i$ from the set \{BeginTransaction, TransactionaFile.Read, TransactionalFile.Write, EndTransactioni\} we would have: \\

Rule 1- for any pair of operations (transactions) $a$ and $b$ their relationship is taken from the set \{$\rightarrow$, $\leftarrow$, $\leftarrow$ and $\rightarrow$\} \\

proof: Follows from defenition. \\

Rule 2) If an operation $a \in OP{T_i}$ has multiple prcedecors (denoted as $OP_{precedesors}$) that their changes overlap each other, then if all of them are from sets other than $OP{T_i}$, $a$ sees the "changes" made by the $op_i \in OP_{predecesors}$ such that $op_i$ has the greatest "commit instant"  . Otherwise, $a$ sees the most recent precedesor in $OP{T_i}$. \\

proof: the first part follows from Def 4 , second derives from Def 4 and second part of Def 1.\\

Rule 3- $\forall op_i \in  OP{T_i}$, $\forall op_j \in  OP{T_j}$, $\forall op_k \in  op{T_k}$ if $op_i \rightarrow op_k$ and $op_k \rightarrow op_j$ then $op_i \rightarrow op_j$ \\

>>>>

A Transaction consists of a sequence of operations. Thus, a transaction can be represented as the \{$OP_1$, $OP_2$, ..., $OP_n$\}. The flow of program should be in a manner that we could have an arbitrary serial order of the members of $T$ (e.g. $T_5,T_6,T_1$,...) or any other order. The serial order should be in such a way that for any given two members of the set $T$, it looks all the operations of one occur before all the operation of the other. In the first case we would say the $T_i$ precedes $T_j$ and in the seconds case $T_j$ preceds $T_i$ and $T_i$ can observe all the effects made by the operations in $T_j$. \\

@@

Def 1- \emph{Set of Operations}: Operations are taken from the set \{readoffset(fildescriptor), writeoffset(filedescriptor), readdata(inode, offset, length), writedata(inode, offset, length)\}. We denote read operations as readoffset and readdata and write operations as writeoffset and writedata.\\

%Def 1- 1-An operation in $T_i$ is said to have "changed" or "affected" some data structure in correspondece to all members of $OP_{T- T_i}$ , when it reflects the changes in the filesystem in such a way that the modification is durable and irreversible. A transaction with such an operation $T_i$ in its operation set is said to have changed the data 2- An operation in $T_i$ is said to have "changed or affeced the data in correspondence to all the operations in $OP_{T_i}$ as soon as the operation is encountered \\

Def 2- \emph{Operations Sharing Resources} 1- A readoffset operation and writeoffset operations are said to be sharing resources if they both operate on the same filedescriptor. 2- A readdata and writedata operations are said to be sharing resources if they both operate on the same inode AND the range of (offset, offset + length) within one overlaps with that of the other.\\

Def 3- A read operation can only see the writes made by write operations sharing reources.\\

Def 4- \emph{Commit Instant}: A transaction $T_i$ is said to "commit" at instant $t_i$, if and only if it reflects its "writes" mabe by write operations in the filesystem or equivalently makes it writes visible to the whole system for all members of $T$ accesing the data at instant $t_j$ such that $t_j > t_i$ . An operation $\in OP_{T_i}$ is said to commit if and only if $T_i$ commits. Commit instant are not splittable, hence it appears to other transactions that all writes are refleted in the file system together.\\

Def 5- \emph{No See In Future}: Assume $T_i$ is an uncomiited transaction and $T_{commited}$ indicates the set of commited transaction. A read operation $b \in OP_{T_i}$ MAY ONLY see the writes(changes) made by write operation $b$ that shares resources with $b$ AND also is subject to one of the following conditions:\\

\hspace{7mm} 1- $a \in OP_{T_j}$ ($j \neq i$)  such that $T_j \in T_{commited}$ \\

\hspace{7mm} 2- $a \in OP_{T_i}$ and $a$ happens before $b$ in the natural order of the transaction. Formally, $a$ has a lower position than that of $b$ in $OP_{T_i}$ = \{$OP_1$, ..., $OP_n$\} \\

Def 7- \emph{Precedence Relationship for Read-Write Operations}: Assume $b$ is a read operation and $a$ is a write operation, and $a$ and $b$ share resource $r_i$, $a$ precedes $b$ ($a \rightarrow b$) if complying with Def 5, $b$ is allowed to see the writes to $r_i$ made by $a$. Otherwise, $b \rightarrow a$. \\

Corrolary- \emph{Read Must See Most Recent Comitted Write}: If $a \in OP{T_i}$ reads the resource $r_i$, in case $a$ has multiple precedessors sharing $r_i$ (denoted as $OP_{precedessors-for-r_i}$), then if all of them are from sets other than $OP_{T_i}$, $a$ sees the writes made by $op_{T_j} \in OP_{precedessors-for-r_i}$ such that $T_j$ has the greatest "commit instant" between all transaction with such operations. Otherwise, $a$ sees the most recent precedessor in $OP_{T_i}$. \\

proof: Consider case 1 that all precedessors are from sets of operations other than that of $a$'s, according to Def 4 all these prcedessors belong to comitted transactions. Assume $T_i$ is accessing data at $t_i$, $t_i$ is neccasarily gretaer than all members of \{$committime_{T_1}, ..., commitime_{T_{i-1}}$\},and hence the writes to $r_i$ made by these commited transaction, have accroding to Def 1 been already reflected in the file system. Since, $T_{i-1}$ has the greatest commit instant its changes have overriiden that of previously commited transactions and thus, $a$ reading $r_i$ from file system sees these changes.\\

In case 2 that there are operations preceding $a$ in $OP_{T_i}$, according to defenition of transaction $r_i$ should be read from the writes made by write operations in $T_i$, and as the last of such write operation overrides the written data to $r_i$ by other operations, $a$ gets to see the most recent precedessor in $T_i$.\\

Def 8- \emph{Precedence Relationship for Other Types of Operations}:  Follwoing rules are ONLY enforced if precedence elationship is not etemined between two transactions by Def 6. 1-If $T_i$ and $T_j$ are two transactions both committed with commit instant of $T_j$ being greater than commit instant of $T_i$ then $\forall op_i \in OP_{T_i}, \forall op_j \in OP_{T_j} op_i \rightarrow op_j$. 2-If two operation belong to the same transaction, then the one that occurs earlier in natural order of transaction precedes the other. 3- If $T_i$ is committed and $T_j$ is not, $\forall op_i \in OP_{T_i}, \forall op_j \in OP_{T_j} op_i \rightarrow op_j$. 4- If neither $T_i$ has commited nor $T_j$, then  $\forall op_i \in OP_{T_i}, \forall op_j \in OP_{T_j} op_i \rightarrow op_j$, if there is either not a precedence relationship between $op_i$ and $op_j$ or it is $op_i \rightarrow op_j$. 5- Otherwise, it means some operations in $T_i$ precede some of thos in $T_j$ and the other way around, for thos pairs not having such relationship it is assigned arbitrarily.\\

Def 9- $\forall$ $op_{T_i}$ $\in OP_{T_i}$ and $\forall op_{T_j}\in OP_{T_j}$ if there is a precedence relationship between $op_{T_i}$ and $op_{T_i}$ if and only if $op_{T_i} \rightarrow op_{T_j}$ then $T_i$ $\rightarrow T_j$ \hspace{8mm} (this defines precedes relationship for members of $T$) \\

%Def 2- An operation in $T_i$ is said to have "used" the data when it has read some data and has done some actions that is potentially irreversible. hence if the data in question was to change, $op$ would have different results. A transaction with such an operation in its operations set is said to have used that data. Any data read and used within a transaction at its commit instant does not count as using. "Use" can just be attributed to the operations actually reading data before commit instant. \\

%Def 4-  $\forall op_i \in OP_{T_i}$, $\forall op_j in OP{T_i}$, $op_i$ precedes $op_j$ if and only if $op_j$ does not use the data changed by $op_i$ OR if it does, then $op_j$ sees the changes made by $op_i$ unless data has been changed by a more recent precedessor. Furthermore  \\

%Def 5- Succeds relationship is defined similarly.  \\

%Def 5- \emph{Precedence Relationship for Operations}: Assume $b$ is a read operation and $a$ is a write operation, and $a$ and $b$ share resource $r_i$, 1- if $b$ see the writes made by the $a$ to $r_i$ and their relationship conforms to Def 4, then $a$ precedes $b$ ($a \rightarrow b$) 2- if $b$ does not see the changes written by $a$, however their relatioship conforms to Def 4,   $r_i$ has been changed by operation $c$ such that $t_{a-commit-instant}<t_{c-commint-instant}$ and $b$ sees the write made by $c$, then $a \rightarrow b$. 3- If $b$ does not see the changes made by $a$ and $r_i$ has been changed by operation $c$ such that 3- Otherwise $b$ precedes $a$. Precedence is only defined for a read and write operation sharing resources.\\

%Def 6- If operation $a$ both precedes and succeds operation $b$ since they do not share any resources so the changes can be made visible, $a$ both precedes and succeds $b$. \\

\subsection{Rules}
%If we denote $\rightarrow$ as the relation precedes between operations and $OP_{T_i}$ as the set of all operation within $T_i$ from the set \{BeginTransaction, TransactionaFile.Read, TransactionalFile.Write, EndTransactioni\} we would have: \\

i%Rule 1- for any pair of operations (transactions) $a$ and $b$ their relationship is taken from the set \{$\rightarrow$, $\leftarrow$, $\leftarrow$ and $\rightarrow$\} \\

%proof: Follows from defenition. \\

%Rule 2) If an operation $a \in OP{T_i}$ has multiple prcedecors (denoted as $OP_{precedesors}$) that their changes overlap each other, then if all of them are from sets other than $OP{T_i}$, $a$ sees the "changes" made by the $op_i \in OP_{predecesors}$ such that $op_i$ has the greatest "commit instant"  . Otherwise, $a$ sees the most recent precedesor in $OP{T_i}$. \\

%proof: the first part follows from Def 4 , second derives from Def 4 and second part of Def 1.\\

Rule 3- $\forall op_i \in  OP{T_i}$, $\forall op_j \in  OP{T_j}$, $\forall op_k \in  OP{T_k}$, if there is a precedence relationship between $op_i$ and $op_k$ sand $op_i \rightarrow op_k$, and if there is a relationship between $op_k$ and $op_j$ and $op_k \rightarrow op_j$, then $op_i \rightarrow op_j$ \\
